---
title: "Lab-12-hw-wolf"
author: "Kyra Kong"
date: '2023-03-01'
output: 
  html_document: 
    keep_md: yes
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Libraries
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(shiny)
library(shinydashboard)
```

## Wolves Example
Let's try to build an app that allows users to explore sexual composition of wolf packs by `pop`. These data are from: Brandell, Ellen E (2021), Serological dataset and R code for: Patterns and processes of pathogen exposure in gray wolves across North America, Dryad, [Dataset](https://doi.org/10.5061/dryad.5hqbzkh51)  

1. Start by loading and exploring the data.
```{r}
wolves <- read_csv("data/wolves_data/wolves_dataset.csv")
```

```{r}
names(wolves)
```

2. For the app, here is a faceted version of what we are looking for. We want the reactive part to be `pop` and you should use `shinydashboard`.
```{r}
wolves %>% 
  filter(sex!="NA") %>% 
  ggplot(aes(x=sex, fill=sex))+
  geom_bar()+
  facet_wrap(~pop)
```
```{r}
library(shiny)

ui <-dashboardPage(
  dashboardHeader(title ="Sexual Composition of Wolves by Population"),
  dashboardSidebar(disable= T),
  dashboardBody(
    fluidRow(
      box(title = "Population Options", width = 3,
        selectInput("x", "Select Population", choices = c("AK.PEN", "BAN.JAS", "BC", "DENALI", "ELLES", "GTNP", "INT.AK", "MEXICAN", "MI", "MT", "N.NWT", "ONT", "SE.AK", "SNF", "SS.NWT", "YNP", "YUCH"), selected = "AK.PEN"),
                plotOutput("plot", width= "600px", height = "400px")
  
))))

server <- function(input, output, session) {
  session$onSessionEnded(stopApp) 
  output$plot <- renderPlot({
    wolves %>% 
      filter(sex!= "NA") %>% 
      filter(pop== input$x) %>% 
      ggplot(aes(x=sex, fill= sex))+
      geom_bar(color= "black", alpha = 0.5)+
      theme_light()
      
  })
  
}

shinyApp(ui, server)
```

