---
title: "Importing Data Frames"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: spacelab
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
---

## Learning Goals
*At the end of this exercise, you will be able to:*    
1. Import .csv files as data frames using `read_csv()`.  
2. Use summary functions to explore the dimensions, structure, and contents of a data frame.  
3. Use the `select()` command of dplyr to sort data frames.  

## Review
At this point, you should have familiarity in RStudio, GitHub, and basic operations in R. You understand how to do arithmetic, assign values to objects, and work with vectors, data matrices, and data frames. If you are confused or need some extra help, please [email me](mailto: jmledford@ucdavis.edu).  

## Load the tidyverse
```{r message=FALSE, warning=FALSE}
library("tidyverse") #install.packages("tidyverse") but do not need to do that if its already installed
```

## Data Frames
For the remainder of the course, we will work exclusively with data frames. Recall that data frames store multiple classes of data. Last time, you were shown how to build data frames using the `data.frame()` command. However, scientists often make their data available as supplementary material associated with a publication. This is excellent scientific practice as it insures repeatability by showing exactly how analyses were performed. As data scientists, we capitalize on this by importing data directly into R.  

## Importing Data
R allows us to import a wide variety of data types. The most common type of file is a .csv file which stands for comma separated values. Spreadsheets are often developed in Excel then saved as .csv files for use in R. There are packages that allow you to open excel files and many other formats directly but .csv is the most common.

An opinionated word about excel. It is fine to use excel for data entry and basic analysis. But, it often adds formatting that makes excel files difficult to work with in any program besides excel. R can read excel files, but I know of no R programmers that routinely use them. Instead they save copies of their excel files as .csv which strips away formatting but makes them easier to use in a variety of programs. We won't work with excel files in BIS 15L, but we will learn to import them.  

To import any file, first make sure that you are in the correct working directory. If you are not in the correct directory, R will not "see" the file.
```{r}
getwd()
```

## Load the data
Here we open a .csv file. Since we are using the tidyverse, we open the file using `read_csv()`. `readr` is included in the tidyverse set of packages. We specify the package and function with the `::` symbol. This becomes important if you have multiple packages loaded that contain functions with the same name.  

In the previous part of the lab, you exported a `.csv` of hot springs data. Let's try to reload that `.csv`.  
```{r}
hot_springs <- readr::read_csv("hsprings_data.csv") #has to be the exactly the same name that you had 
```

Use the `str()` function to get an idea of the data structure of `hot_springs`.  
```{r}
str(hot_springs)
```

```{r}
spec(hot_springs)
```

What is the class of the scientist column? Change it to factor and then show the levels of that factor.  
```{r}
class(hot_springs$scientist)
```


```{r}
hot_springs$scientist <- as.factor(hot_springs$scientist) # treat data as a factor, changing the class
class(hot_springs$scientist)
```

```{r}
levels(hot_springs$scientist)
```

## Practice
1. In your lab 3 folder there is another folder titled `data`. Inside the `data` folder there is a `.csv` titled `Gaeta_etal_CLC_data.csv`. Open this data and store them as an object called `fish`.  

The data are from: Gaeta J., G. Sass, S. Carpenter. 2012. Biocomplexity at North Temperate Lakes LTER: Coordinated Field Studies: Large Mouth Bass Growth 2006. Environmental Data Initiative.  [link](https://portal.edirepository.org/nis/mapbrowse?scope=knb-lter-ntl&identifier=267)

```{r}
getwd() #can copy this infront of data/ ex: fish <- readr:read_csv (/Users/kyrkon38/Desktop/BIS15W2023_kkong-main/lab3/data/) # in windows switch the slashes from back slashes to front slashes
```

```{r}
fish <-readr::read_csv("data/Gaeta_etal_CLC_data.csv") #insert data/ in quotations (or a specific folder within the working directory). specify the path
```

2. What is the structure of these data?
```{r}
str(fish) # 4033 obs and 6 variables
```

Notice that when the data are imported, you are presented with a message that tells you how R interpreted the column classes. This is also where error messages will appear if there are problems. (it can make mistakes, so you can change the classes)

## Summary functions
Once data have been uploaded, you may want to get an idea of its structure, contents, and dimensions. I routinely run one or more of these commands when data are first imported.  

We can summarize our data frame with the`summary()` function.  
```{r}
summary(fish) #  for numeric can do numbers like mean, median, min, max and not for characters and it removes NAs automatically

```

`glimpse()` is another useful summary function.
```{r}
glimpse(fish)
```

`nrow()` gives the numbers of rows.
```{r}
nrow(fish) #the number of rows or observations
```

`ncol` gives the number of columns.
```{r}
ncol(fish) #the number of columns or variables
```

`dim()` gives the dimensions.
```{r}
dim(fish) #total dimensions
```

`names` gives the column names.
```{r}
names(fish) #column names (most useful for us)
```

`head()` prints the first n rows of the data frame.
```{r}
head(fish, n = 10)
```

`tail()` prints the last n rows of the data frame.
```{r}
tail(fish, n = 10)
```

`table()` is useful when you have a limited number of categorical variables. It produces fast counts of the number of observations in a variable. We will come back to this later... 
```{r}
table(fish$lakeid) #make a table from data set fish and create lake id
```

We can also click on the `fish` data frame in the Environment tab or type View(fish).
```{r}
View(fish)
```

## Filter
Filter is a way of pulling out observations that meet specific criteria in a variable. We will work a lot more with this in the next lab.  
```{r}
little_fish <- filter(fish, length<=100) #take out fish with lengths less than or equal to 100
little_fish
```

## Practice
1. Load the data `mammal_lifehistories_v2.csv` and place it into a new object called `mammals`.
```{r}
mammals <- readr::read_csv("data/mammal_lifehistories_v2.csv")
```

2. Provide the dimensions of the data frame.
```{r}
dim(mammals)
```

3. Check the column names in the data frame. 
```{r}
names(mammals)
```

4. Use `str()` to show the structure of the data frame and its individual columns; compare this to `glimpse()`. 
```{r}
str(mammals)
```

```{r}
glimpse(mammals) #makes most sense
```

5. . Try the `table()` command to produce counts of mammal order, family, and genus.  
```{r}
table(mammals$order) # can only do one column at a time
```
```{r}
table(mammals$family)
```
```{r}
table(mammals$Genus)
```
```{r}
names(mammals)
```


## Wrap-up
Please review the learning goals and be sure to use the code here as a reference when completing the homework.  
-->[Home](https://jmledford3115.github.io/datascibiol/)