---
title: "Lab 11 Intro"
date: "`r Sys.Date()`"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library("jpeg")
```

## Seating & Set-up
1. Please make sure that you sit next to your group members for the rest of the quarter.
2. Please set-up your computer as normal.

## Warm-up
For this warm-up, please use the `heartrate` data. Be sure to label your plots!
```{r}
library(tidyverse)
library(here)
library(janitor)
library(naniar)
```

```{r}
heartrate <- read.csv(here("data", "heartrate.csv"))
```
```{r}
glimpse(heartrate)
```

```{r}
heartrate
```

1. Make a plot that compares the heartrate of patients for drug a. (Relatively easier to do with an X and Y)
```{r}
heartrate %>% 
  ggplot(aes(x=patient, y= a, fill= patient)) +
  geom_col() +
  labs(title = "Comparison of Heartrate for Patients Across Drug A",
       x= "Patient Names", y = "Heartrate")+
  theme_
```

2. Make a plot that compares heartrate (as a range) for each drug. (PIVOTS AGAIN)
How do you make it transparent?
```{r}
heartrate_longer <-heartrate %>%
  pivot_longer(-patient, # patient not pivot
               names_to= "drug", # new column
               values_to = "heartrate" # values are moved
               )
  
```

```{r}
heartrate_longer %>% 
  ggplot(aes(x=drug, y= heartrate, fill = drug)) +
  geom_boxplot(alpha=0.8)+ ## adjusts transparency
  labs(title= "Range of Heartrates per Drug",
       x= "Treatment Type", y = "Heartrate")+
  theme(plot.title = element_text(size = rel(1), hjust = 0.5))+
  theme_bw()
```

