---
title: "Lab 9 Intro"
date: "`r Sys.Date()`"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library("jpeg")
```

## Seating
1. If you already have a group, it will be helpful to start sitting next to each other.
2. If you don't have a group, hang tight we are working on it.
3. Please setup your computer as normal for today.

## Warm-up
1. In the data folder there is an epidemiology data set on an outbreak of malaria.
2. `rdt` refers to rapid diagnostic test and they are identified here by age group.
3. Make the data tidy and store them as a new object.
4. Which district had the highest *total* number of cases on July 30, 2020?

```{r}
library(tidyverse)
library(here)
```
```{r}
getwd()
```
# what the data set is 
```{r}
malaria <- read.csv(here("data", "malaria.csv"))
malaria
```
# making data tidy 
```{r}
malaria_longer <-malaria %>% 
  pivot_longer(-c("location_name", "data_date", "submitted_date", "Province", "District", "malaria_tot", "newid"),
               names_to = c("malaria_rdt"),
               values_to = "tests_ran")
malaria_longer

  
```
```{r}

```

homework for the next weeks:
shopping for datasets
canvas
