---
title: "lab 4 warmup"
output: html_document
date: "2023-01-19"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

```{r}
library(tidyverse)

```

```{r}
library(palmerpenguins)
```
## identify the dimensions
```{r}
penguins ## penguins is already in the library so you can just type penguins
dim(penguins)
glimpse(penguins)
```
Dimensions are 344 x 8 or 344 instances of 8 variables

## identify names of variables
```{r}
names(penguins) ## constantly use this function: important for spelling purposes
```


The names of the variables are species, island, bill length (mm), bill depth (mm), flipper length (mm), body mass (g), sex, and year
## identify how many individuals are sampled on each island
```{r}
table(penguins$island) # in penguins make a table for island, does not work well for tables that aren't categorical 
```

168 penguins on Biscoe, 124 on Dream, and 52 at Torgersen.

## what is the mean body mass for all the individuals in the penguins dataset

```{r}
mean(penguins$body_mass_g, na.rm =T) ## there is NA in the dataset, so removal is necessary, doesn't work for a class of data that is a categorical (classes for variables in a dataset MATTERS)
```
