---
title: "Lab 7 Intro"
date: "`r Sys.Date()`"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library("jpeg")
```

## Setup
1. Login to the lab computer (please don't use your personal computer).  
2. Navigate to github.com and login.   
2. Use GitHub desktop to clone your repository to the desktop.   
5. Copy the class repository to the desktop (https://github.com/jmledford3115/datascibiol).  
6. Copy the files for today's lab from the class repository and paste them into **your** repository.  
7. Open today's lab in RStudio.  

## Review from last time
### *With a partner, discuss the following questions*
1. How did the midterm go? Did you learn anything while doing the exam?

Relatively ok

2. Where did you get stuck?

at the part where you had to summarize the diversity of species, because you could use like 4-5 more lines

3. What do we need to practice?

Memorization of particular functions (so that there's no problem). Practicing the homework and labs again.
### Warm-up
1. Please install the package `naniar`.

```{r}
#install.packages("naniar")
```

```{r}
library(naniar) 
```

#key to the midterm is posted